Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-11-18 special-issue-ci/Response to reviewers.tex

In a nutshell, what we are saying is that, as is proven in the paper,
a JavaScript application, running in the browser, can gather in
seconds dozens of users. That would be, in practice, impossible with
other kind of native clients, who would have to download and maybe
compile the application, losing many seconds or minutes for the
optimization problem that has to be solved.

These facts, along the explanations, have now been added to the first
paragraph of the {\em Description of the system} section, as well as
to the introduction.

\begin{quote}
The paper organization is not clear. Some comments about this:
- The title "Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed
Evolutionary Computation" seems to suggest that the volunteers run in the cloud, but that is not the case:
the server run in the cloud and the volunteers do in the browsers of
common PCs and laptops.
\end{quote}

That is correct. We just used {\em cloud} in a relatively informal
sense because they are available as in a cloud for running the
experiment. That is, from the point of view of the person running the
experiment, there is a {\em cloud} out there running your experiments,
which is actually run by volunteers in their browsers. We went to the
original acceptation of ``cloud'' as a ``network'' more than the more
precise and technical acceptation used today. %PABLO: no existe aception, lo he cambiado a acceptation (segun WordReference)

We have added this explanation to the first paragraph in the
conclusions; we hope this makes our intention clearer, but we can also
eliminate that part of the title if the reviewer thinks it is
necessary.

@@

\begin{quote}
- According the last paragraph of the introduction, ``Then, a detailed description of the proposed system is
presented in Section III, followed by the experiments and obtained results in Section IV.", but Section IV
does not only contains experimental result but also the full description of the second version of the
architecture.
\end{quote}

You are absolutely right. We have corrected that paragraph to include
the actual contents of the paper.

\begin{quote}
Summarizing the review, the authors should provide a convincing example showing the advantages of
their proposal. As presented, the paper is a proof of concept of a system that seems difficult to be used
by the researchers needing high computational resources to run an evolutionary algorithm to solve a realworld
problem.
\end{quote}

We have added a new experiment, a more complex and taxing problem
using a web-worker architecture, in order to prove how this can be
used in real-life problems. In fact, our architecture {\sf NodIO} is
not an exclusively browser-based one. Its pool-based architecture allows
any other kind of client, including local ones, to be incorporated to
the experiment, with the lightweight overhead of doing asynchronous
HTTP requests, which being local are indeed negligible.

@@

Our framework, in fact, does not prevent any other kind of
experiments. If there is a very efficient local algorithm, it only
needs to be able to do HTTP requests to work alongside a {\em
  volunteer cloud}. What our framework ensures is that all work done
by volunteers can be integrated with local, cloud or other kind of
resources. This explanation has been also added as a paragraph in the
Introduction to the paper.

We would like to thank this reviewer for the helpful comments and
suggestion made. We expect to have addressed them properly, but in any
case are grateful because it has allowed us to improve the quality of
our work.

### 2015-11-18 special-issue-ci/Response to reviewers.tex

\noindent
Yours sincerely,\\
The authors.

\section{Response to Reviewer \#1}

\begin{quote}
The main contribution of this paper is the proposal of a Javascript based distributed system intended to
run evolutionary algorithms according to the volunteer computing scheme. The idea is that user can
contribute to the computation by using only a browser, without needing to install any plugin nor a Java
virtual machine.
The volunteer computing model appeared some years ago and it has been used in many fields. The
model is effective to solve problems demanding high computational capabilities, and at this point is where
this paper proposal introduces some concerns. The choice of Javascript code running on browsers has
some advantages, as is pointed out in the paper, but some important drawbacks. Given that the
motivation of these kinds of distributed models is to solve complex problems, using Javascript doesn\'{}t
appear a right choice. It\'{}s true that is a very commonly used programming language, but in the context of
Web applications, not in high performance ones. Apart from the fact of being an interpreted language
designed for the Web, it is difficult to think on a computer scientist writing a complex math code in
Javascrip. If the authors would have used as a case study a complex optimization problem then they
could have provided numerical data to sustain their claims; however, the selected problem (which should
be formulated in the paper) seems a rather academic one.
\end{quote}

We thank the reviewer for the comment. The only reason why JavaScript
was chosen was because it is the only language that is native to the
browser, allowing to carry out precisely this kind of experiments
where the user is invited to click on an URL and start giving CPU to
it. Although there are many other volunteer frameworks, written in
other languages, that a priori would offer a higher performance, the
key is that they would {\em while they are running}, that is, as soon
as it has been installed, the user has decided to fire it up, and it
does not affect the performance of the rest of his system, which is a % se confundia con el desempeño del todo el sistema
concern in this kind of systems. We have performed speed tests on
JavaScript and other compiled languages resulting that, in fact,
JavaScript uses a JIT compiler that provides a good performance; we
have published these tests as an open access technical report.  % quizás indicar la URL aquí

@@

In a nutshell, what we are saying is that, as is proven in the paper,
a JavaScript application, running in the browser, can gather in
seconds dozens of users. That would be, in practice, impossible with
other kind of native clients, who would have to download and maybe
compile the application, losing many seconds or minutes for the
optimization problem that has to be solved.

These facts, along the explanations, have now been added to the first
paragraph of the {\em Description of the system} section, as well as
to the introduction.

\begin{quote}
The paper organization is not clear. Some comments about this:
- The title "Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed
Evolutionary Computation" seems to suggest that the volunteers run in the cloud, but that is not the case:
the server run in the cloud and the volunteers do in the browsers of
common PCs and laptops.
\end{quote}

That is correct. We just used the term {\em cloud} in a relatively informal
sense because they are available as in a cloud for running the
experiment. That is, from the point of view of the person running the
experiment, there is a {\em cloud} out there running your experiments,
which is actually run by volunteers in their browsers. We went to the
original aception of ``cloud'' as a ``network'' more than the more
precise and technical aception used today.

@@

\begin{quote}
- According the last paragraph of the introduction, ``Then, a detailed description of the proposed system is
presented in Section III, followed by the experiments and obtained results in Section IV.", but Section IV
does not only contains experimental result but also the full description of the second version of the
architecture.
\end{quote}

You are absolutely right. We have corrected that paragraph to include
the actual contents of the paper.

\begin{quote}
Summarizing the review, the authors should provide a convincing example showing the advantages of
their proposal. As presented, the paper is a proof of concept of a system that seems difficult to be used
by the researchers needing high computational resources to run an evolutionary algorithm to solve a realworld
problem.
\end{quote}

We have added a new experiment, a more complex and taxing problem
using a web-worker architecture, in order to prove how this can be
used in real-life problems. In fact, our architecture {\sf NodIO} is
not an exclusively browser-based one. Its a pool-based architeture that allows
any other kind of client, including local ones, to be incorporated to
the experiment, with the lightweight overhead of doing asynchronous
HTTP requests, which being local are indeed negligible.

@@

\end{quote}

We have inserted a paragraph at the end of the section that describes
the experimental setup describing what are the challenges that this
kind of voluntary framework must overcome. Another paragraph analyzing
the security from the point of view of the framework used has been
also added. A new paragraph on the security challenges has also been
inserted, third paragraph in this section, to clarify our integral and
pragmatic approach to security, which so far has worked perfectly.

\begin{quote}
-  an experiment with a more difficult problem (please use the well-known benchmark problems
described in literature).
\end{quote}

An experiment based on the well known Rastrigin's function has been
performed and analyzed.

@@

\begin{quote}
-  A comparison in similar conditions of the two implementations of the architecture (the algorithms
used in the experiments are slightly different, but this could lead to dramatic changes in
performance, and consequently, invalidate the comparison performed in the paper). This would
enable a measurement of the actual improvement in performance obtained by the use of web
workers (not that generated due to a different algorithm)
\end{quote}

This has been included as the last subsection of the paper. The new
experiment has a constant population, instead of using  population of random size.

@@

\section{Response to Reviewer \#3}

\begin{quote}
1. Contribution: The paper presents the idea of a web browser based distributed evolutionary computing.
the system called NodIO consists of a REST server (coded in Node.js) and clients performing
evolutionary algorithm coded in JavaScript and embedded in a web page. Although the idea is not
entirely new, the authors show how Web Workers (introduced in HTML5) and the mechanism of
automatic restart of computations on the client's side can improve the performance of the system. The
state of the art is comprehensive and provides good introduction into the subject.

2. Relevance: In my opinion the system described in the paper is relevant to "Computational Intelligence
Software" special issue. Thanks to both development of Internet technology and continuous increase in
computing power of electronic devices, including mobile devices, the web browser based distributed
computing can be a cheap alternative for grid and high-performance computing systems that require
high capital expenditures and consume a lot of energy. Evolutionary computing system based on the
architecture presented inthe paper can be potentially used for solving many different time consuming
problems, however, the authors limited their experiments to a simple "Trap" function that can be easily
solved on a desktop machine, what may not be convincing for a reader looking for an effective
optimization tool.
\end{quote}

It is true that that function can be easily solved with any desktop
machine, but we can still do that {\em and} use a volunteer
computation framework, which will add performance to the desktop
machine. Our intention was more to prove what kind of performance can
be expected and how this performance is distributed, although we have
also proved in the last section how the performance achieved by a
volunteer-based computer system can, in fact, be better than a desktop, even
a powerful one. This has also been clarified in the introduction and
conclusions.

At any rate, we have also performed a new experiment based on the
Rastrigin's function, which in fact cannot be solved in a few minutes
(or seconds, as in the case of Trap) by a desktop system.

@@

We have extended the abstract stressing our main points, that were
first to test the validity and measure the performance of our
framework NodIO and then to use the measures as first steps to a model
of volunteer uses.

\begin{quote}
Comments to Authors:

1) I like the state of the art which includes comprehensive overview of the most important papers related
to the subject presented in the paper. I also like the description of the system, but I cannot fully agree
with its characteristic provided on page 4. Namely:

 - Heterogeneity - it is true that almost all devices can participate in the web browser based
 distributed computing system, providing that a user did not disable Java Script in its browser.
 However if the authors plan to use HTML5 Web Workers only modern web browsers can
 participate in such computing system, and some devices (e.g. based on Android 2.2-4.3 or
 Windows Phone) are also excluded.

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Adds to conclusions refs #40 closes #2

Although we think that, in general, the results presented in this
paper are independent of the problem chosen, it is true that even if
it is a difficult problem for evolutionary algorithms, it takes a few
seconds with the right settings to solve. This makes it difficult also
for this kind of asynchronous framework, but it would be interesting
to check the performance for more computationally demanding
tasks. Besides, the two versions of the framework include as an
algorithmic variant using random population size. We do not think this
has had a big influence on the results, but it would be interesting to
measure exactly what this influence has been. Adding a bit of more
control to the user might contribute also to gamification. Right now
the user has only two Web Workers. It is a matter of a single click to
open more tabs in the browser, giving more Web Workers, but it would be
interesting to put everything in a single page and under our control,
to check how often this happens and under which conditions.

>>>>

Although we think that, in general, the results presented in this
paper are independent of the problem chosen, it is true that even if
it is a difficult problem for evolutionary algorithms, it takes a few
seconds with the right settings to solve. This makes it difficult also
for this kind of asynchronous framework, but it would be interesting
to check the performance for other computationally demanding
tasks, especially ones that cover the middle ground between a problem
finishing in seconds or minutes (Trap) and weeks (Rastrigin F15). 
Besides, the two versions of the framework include as an
algorithmic variant using random population size. We do not think this
has had a big influence on the results and in fact this is not noticed
for F15, which is the second version tested but it would be interesting to
measure exactly what this influence has been. In general there are
many issues with the evolutionary algorithm implementation itself,
including using different, or adaptive, policies for inserting and
sending individuals to the pool as was done in \cite{araujo2008mam},
using different policies for population initialization, and also the
incorporation of high-speed local resources to the pool to check what
would be the real influence of the volunteer pool to the final
performance.

Another area is how to enhance the number and quality of
volunteers. For instance, adding a bit of more
control to the user might contribute also to gamification. Right now
the user has only two Web Workers. It is a matter of a single click to
open more tabs in the browser, giving more Web Workers, but it would be
interesting to put everything in a single page and under our control,
to check how often this happens and under which conditions.
